---
title: 'Chapter 2: Examples'
output: 
    html_document:
        toc: true
---

___
# Getting the data
- Import the library and the dataset 
```{r 'library import 1'}
library(AppliedPredictiveModeling)
data(FuelEconomy)
```

- Check the size of the different data frames
```{r 'comparing dimensions of the cars data'}
dimensions <- sapply(list(cars2010, cars2011), dim)
dimnames(dimensions) <- list(c("rows", "columns"), c("cars2010", "cars2011"))
dimensions
```

- Check for missing values
```{r 'checking for missing values'}
countNa <- function(x){
    sum(is.na(x))
}

sapply(list(cars2010, cars2011), apply, 2, countNa)
```


## Formatting the data
- Sort by engine displacement
```{r 'sort by engine displacement'}
cars2010 <- cars2010[order(cars2010$EngDispl), ]
cars2011 <- cars2011[order(cars2011$EngDispl), ]


# print the first rows to confirm
lapply(list(cars2010[ , 1:5], cars2011[ , 1:5]), head)
```

- Combine data into one data frame
```{r 'combine data for plotting'}
cars2010a <- cars2010
cars2010a$Year <- "2010 Model Year"
cars2011a <- cars2011
cars2011a$Year <- "2011 Model Year"

plotData <- rbind(cars2010a, cars2011a)
```

___
# Exploring the data
## Plotting fuel economy against engine displacement
```{r 'plot1', fig.align='center'}
library(ggplot2)
ggplot(plotData, aes(x = FE, y = EngDispl)) +
    geom_point(alpha = 1/2) +
    labs(x = "Engine Displacement", y = "Fuel Efficiency (MPG)") +
    facet_grid(. ~ Year)
        
```

___
# Fitting models to data
## Fit a single linear model using 10-fold CV with the 'caret' package
```{r 'model fitting lm'}

library(caret)
set.seed(1)
lmOneFit <- train(FE ~ EngDispl, 
                    data = cars2010, 
                    method = "lm",
                    trControl = trainControl(method = "cv"))
lmOneFit
```

- Define plotting functions
```{r 'defining plots with ggplot2'}

library(ggplot2)
library(gridExtra)

plotModel <- function(model, data, title){
    data$modelFit <- predict(model, data)
    
    plotFit <- ggplot(data) +
        geom_point(aes(x = EngDispl, y = FE), alpha = 1/2) +
        geom_line(aes(x = EngDispl, y = modelFit, color = 'red')) + 
        guides(color = FALSE) +
        ggtitle(title) +
        labs(x = "Engine Displacement", y = "Fuel Efficiency (MPG)") +
        facet_grid(. ~ Year)

    plotCompare <- ggplot(data, aes(x = FE, y = modelFit)) +
        geom_point(alpha = 1/2) +
        geom_abline(slope = 1, linetype = 'dashed', color = 'grey') + 
        xlim(10, 70) +
        ylim(10, 70) + 
        labs(x = "Observed", y = "Predicted") +
        facet_grid(. ~ Year)

grid.arrange(plotFit, plotCompare, nrow = 2) 
}

```

- Plot linear model
```{r 'plot linear', fig.align='center', fig.height=7, fig.width=7}
    plotModel(lmOneFit, plotData, 'Linear Model')
```

## Fit a quadratic model
```{r 'fitting quadratic model'}
cars2010$ED2 <- cars2010$EngDispl^2
cars2011$ED2 <- cars2011$EngDispl^2
plotData$ED2 <- plotData$EngDispl^2

set.seed(1)
lmTwoFit <- train(FE ~ EngDispl + ED2,
                    data = cars2010,
                    method = 'lm',
                    trControl = trainControl(method = 'cv'))
lmTwoFit

```

- Plot quadratic model
```{r 'plot quadratic', fig.align='center', fig.height=7, fig.width=7}
    plotModel(lmTwoFit, plotData, 'Quadratic Model')
```

## Fit a MARS model (earth package)
```{r 'fit MARS', fig.align='center'}
library(earth)
set.seed(1)
marsFit <-train(FE ~ EngDispl,
                data = cars2010,
                method = "earth",
                tuneLength = 15,
                trControl = trainControl(method = "cv"))
marsFit

# Plot tuning parameter profile
plot(marsFit)
```

- Plot MARS model
```{r 'plot MARS', fig.align='center', fig.height=7, fig.width=7}
    plotModel(marsFit, plotData, 'MARS Model')
```

___
# Testing model performance
## Get test set performance values via caret's postResample function
```{r 'test performances'}
## Predict the test set data
cars2011$lm1  <- predict(lmOneFit,  cars2011)
cars2011$lm2  <- predict(lmTwoFit,  cars2011)
cars2011$mars <- predict(marsFit, cars2011)

postResample(pred = cars2011$lm1,  obs = cars2011$FE)
postResample(pred = cars2011$lm2,  obs = cars2011$FE)
postResample(pred = cars2011$mars, obs = cars2011$FE)
```

