---
title: 'Chapter 3: Data Pre-processing'
output: html_document
---

___
#### Data Import
```{r 'Case Study'}
library(AppliedPredictiveModeling)
data(segmentationOriginal)

## Retain the original training set
segTrain <- subset(segmentationOriginal, Case == "Train")

## Remove the first three columns (identifier columns)
segTrainX <- segTrain[, -(1:3)]
segTrainClass <- segTrain$Class
```

___
#### Data Transformations for Individual Predictors

###### Skewness:
A rule of thumb for checking skewness of data is the ratio between the maximum and the minimum values such that $\frac{max}{min} > 20$ is a sign of skewness. 
The default measure of skewness implemented in the 'e1071' package is calculated by the formula:
$$
skewness = 
\frac{
    \frac{1}{n} \sum_{i=1}^{n} (x_{i} - \bar{x})^3
}{
    [
        \frac{1}{n-1} \sum_{i=1}^{n} (x_{i} - \bar{x})^2
    ]^\frac{3}{2}
}
$$.

Testing a single predictor for skewness:
```{r 'Measuring skewness'}
## The column VarIntenCh3 measures the standard deviation of the intensity
## of the pixels in the actin filaments
## Getting the ratio of max vs. min values of VarIntenCh3
max(segTrainX$VarIntenCh3)/min(segTrainX$VarIntenCh3)

## Checking skewness of VarIntenCh3
library(e1071)
skewness(segTrainX$VarIntenCh3)
```

###### Dealing with skewness:
Some data might be skewed in terms of natural units, but normally distributed on a logarithmic scale.
```{r 'Plot natural vs log transformation', fig.align='center', fig.height=7, fig.width=7}
library(ggplot2)
## Plotting original and log-transformed variable VarIntenCh3
plotNatural <- ggplot(segTrainX, aes(x = VarIntenCh3)) +
    geom_histogram(color = "gray") +
    labs(x = "Natural Units")

plotLog <- ggplot(segTrainX, aes(x = log(VarIntenCh3))) +
    geom_histogram(color = "gray") +
    labs(x = "Log Units")

library(gridExtra)
## Suppress stat_bin message and position_stack (rounding?) warning
suppressWarnings(suppressMessages(grid.arrange(plotNatural, plotLog, ncol = 2)))

```

A more generalized approach as proposed by Box and Cox (1964) estimates a parameter $\lambda$ using MLE and transforms the data according to the formula:
$$
\begin{aligned}
    x =
    \begin{cases}
        \frac{x^{\lambda} - 1}{\lambda} & if \lambda \neq 0 \\
        log(x) & if \lambda = 0 \\
    \end{cases}
\end{aligned}
$$
```{r 'BoxCox transformation'}
## Use caret's preProcess function to transform for skewness
library(caret)
segPP <- preProcess(segTrainX, method = "BoxCox")

## Apply the transformations and save in new data.frame
segTrainTrans <- predict(segPP, segTrainX)
```

```{r 'BoxCox results 1'}
## Results for a predictor VarIntenCh3
segPP$bc$VarIntenCh3
```

With $\lambda = 0$ the transformation will be the natural logarithm like we did before.
```{r 'Plot natural vs BoxCox transformation 1', fig.align='center', fig.height=7, fig.width=7}
## Plotting original and BoxCox-transformed variable VarIntenCh3
plotNatural <- ggplot(segTrainX, aes(x = VarIntenCh3)) +
    geom_histogram(color = "gray") +
    labs(x = "Natural Units")

plotLog <- ggplot(segTrainTrans, aes(x = VarIntenCh3)) +
    geom_histogram(color = "gray") +
    labs(x = "Transformed")

## Suppress stat_bin message and position_stack (rounding?) warning
suppressWarnings(suppressMessages(grid.arrange(plotNatural, plotLog, ncol = 2)))
```

```{r 'BoxCox results 2'}
## Results for a predictor PerimCh1
segPP$bc$PerimCh1
```

The 'BoxCox'transformation with $\lambda = -1.1$ results in the following distribution.
```{r 'Plot natural vs BoxCox transformation 2', fig.align='center', fig.height=7, fig.width=7}
## Plotting original and BoxCox-transformed variable PerimCh1
plotNatural <- ggplot(segTrainX, aes(x = PerimCh1)) +
    geom_histogram(color = "gray") +
    labs(x = "Natural Units")

plotLog <- ggplot(segTrainTrans, aes(x = PerimCh1)) +
    geom_histogram(color = "gray") +
    labs(x = "Transformed")

## Suppress stat_bin message and position_stack (rounding?) warning
suppressWarnings(suppressMessages(grid.arrange(plotNatural, plotLog, ncol = 2)))
```

___
#### Data Transformations for Multiple Predictors
##### Two predictors
Data transformation for two predictors using the 'prcomp' function from the 'stats' package
```{r 'Data Transformations using prcomp', fig.align='center', fig.height=7, fig.width=8}

## R's prcomp is used to conduct PCA
pr <- prcomp(~ AvgIntenCh1 + EntropyIntenCh1, 
             data = segTrainTrans, 
             scale. = TRUE)

## Scatterplot of original and extracted predictors
plotOriginal <- ggplot(segTrainTrans, aes(x = AvgIntenCh1, y = EntropyIntenCh1)) +
    geom_point(aes(color = segTrain$Class), alpha = 1/2) +
    guides(color = FALSE) +
    labs(title = "Original", x = "Channel 1 Fiber Width", y = "Intensity Entropy Channel 1")

plotTrans <- ggplot(as.data.frame(pr$x), aes(x = PC1, y = PC2)) +
    geom_point(aes(color = segTrain$Class), alpha = 1/2) +
    labs(title = "Extracted",x = "Principle Component #1", y = "Principle Component #2") +
    xlim(c(-4,4)) +
    ylim(c(-4,4)) + 
    theme(legend.position = "top", legend.title = element_blank())

## Suppress stat_bin message and position_stack (rounding?) warning
suppressWarnings(suppressMessages(grid.arrange(plotOriginal, plotTrans, ncol = 2)))
```
Note: The axes in the book for the first plot are flipped and incorrectly labeled.

##### Entire dataset
Zero variance predictors need to be removed, because PCA uses variance to determine the components.
```{r 'Find and remove zero variance predictors'}
## Check for zero variance predictors
isZV <- apply(segTrainX, 2, function(x) length(unique(x)) == 1)

## Count
sum(isZV)

## Show
isZV[isZV == TRUE]

## Remove
segTrainX <- segTrainX[ , !isZV]
```

Data Transformation of the entire dataset using the interface offered by the 'caret' package.
```{r 'Data transformation with caret'}
## Get the pre-process transformation
segPP <- preProcess(segTrainX, method = c("BoxCox", "center", "scale"))

## Apply the transformation
segTrainTrans <- predict(segPP, segTrainX)

## PCA using prcomp
segPCA <- prcomp(segTrainTrans, center = TRUE, scale. = TRUE)
## PCA using caret
segPCA2 <- preProcess(segTrainTrans, method = c("pca"))

## The dataset of principle components
segPCATrans <- as.data.frame(segPCA$x)
segPCATrans2 <- predict(segPCA2, segTrainTrans)

## Comparing first three components
lapply(list(segPCATrans[ , 1:3], segPCATrans2[ , 1:3]), head)
```

Plot matrix of the first three components
```{r 'Plotting 3 components', fig.align='center', fig.height=7, fig.width=7}
## Plot matrix for the first 3 components
library(ggplot2)
library(GGally)
ggpairs(cbind(segPCATrans, segTrainClass),
        columns = 1:3,
        upper = list(continuous = "cor"),
        color = "segTrainClass",
        alpha = 1/2, 
        axisLabels = "internal"
        )
```

Deriving the channels that are associated with the original predictors.
```{r 'Derive channel variable'}
## Get predictor channels 
vars <- rownames(segPCA$rotation)
channel <- rep(NA, length(vars))
channel[grepl("Ch1$", vars)] <- "CH 1"
channel[grepl("Ch2$", vars)] <- "CH 2"
channel[grepl("Ch3$", vars)] <- "CH 3"
channel[grepl("Ch4$", vars)] <- "CH 4"
channel <- as.factor(channel)

## Number of predictors not associated with any channel
sum(is.na(channel))
```

Plot the original predictors' loadings for the first three components and color by the corresponding channels.
```{r 'Rotations', fig.align='center', fig.height=7, fig.width=7}
## Get rotations
segRot <- as.data.frame(segPCA$rotation[ , 1:3])

## Add channel 
segRot$channel <- channel
## Remove incomplete cases
segRot <- segRot[complete.cases(segRot), ]

## Plot matrix
matrixPlot <- ggpairs(segRot,
        columns = 1:3,
        color = "channel",
        alpha = 1/2, 
        axisLabels = "none"
        )
matrixPlot
```


