---
title: "Decorators in R"
author: "Niko Steinhoff"
date: "18 Jan 2015"
output: html_document
---

A few examples of how Python-style decorators (@decorator) can be done in *R* in order to add functionality to existing functions.

#### A decorator function for logging the function's arguments and return value.
```{r 'Logging decorator'}
makeLogged <- function(func){
    function(...){
        returnValue <- func(...)
        
        args <- list(...)
        
        print("Arguments:")
        for(i in 1:length(args)){
            print(paste(names(args)[i], args[i], sep = " = "))
        }
        
        print(paste("Returns", returnValue, sep = ": "))
        
        force(returnValue)
    }   
}
loggedMean <- makeLogged(mean)
loggedSum <- makeLogged(sum)
```

The original function that was passed to the decorator is saved in the environment of the resulting decorated version.
```{r 'Looking at the environment'}
env <- environment(loggedMean)
ls(env)
env$func

environment(loggedSum)$func
```

The decorated function is called exactly like the original and returns the same value.
```{r 'Calling loggedMean'}
mean(c(-10000,1:10, 99999), trim = 0.1)
loggedMean(c(-10000,1:10, 99999), trim = 0.1)
```

The decorated function matches named arguments.
```{r 'Calling loggedSum'}
sum(na.rm = TRUE, 1, 2, three = 3, c(10,NA,12), onehundred = 100)
loggedSum(na.rm = TRUE, 1, 2, three = 3, c(10,NA,12), onehundred = 100)
```
___

#### A decorator for making any function cached
```{r 'Caching decorator'}
makeCached <- function(func){
    cachedReturnValue <- NULL
    lastArgs <- NULL
    function(...){
        needToEvaluate <- FALSE
        if(is.null(cachedReturnValue) || is.null(lastArgs)){
            needToEvaluate <- TRUE
            print("First time")
        } else if(!identical(lastArgs, list(...))){
            needToEvaluate <- TRUE
            print("Different arguments")
        }
        
        if(needToEvaluate){
            print("Evaluating function")
            returnValue <- func(...)
                 
            ## '<<-' double assignment operator needed to search in parent environment
            cachedReturnValue <<- returnValue
            lastArgs <<- list(...)
        } else{
            print("Retrieving cached value")
            returnValue <- cachedReturnValue
        }
   
        force(returnValue)
    }
}
```

Calling the cached function:
```{r 'Calling cached function'}
cachedMean <- makeCached(mean)
cachedMean(c(1,2,3))
cachedMean(c(1,2,3))
cachedMean(c(1,2,3,5))
cachedMean(c(1,2,3,5))
```

Combining decorators: 
The second time around the loggedSum function doesn't get called, hence no logging.
```{r 'Combining operators'}
cachedLoggedSum <- makeCached(loggedSum)
cachedLoggedSum(na.rm = TRUE, 1, 2, three = 3, c(10,NA,12), onehundred = 100)
cachedLoggedSum(na.rm = TRUE, 1, 2, three = 3, c(10,NA,12), onehundred = 100)
```

Using the cacing decorator for more complex return types:
```{r 'More complex functions'}
data(cars)

lm(dist ~ speed, data = cars)

cachedLm <- makeCached(lm)
cachedLm(dist ~ speed, data = cars)
cachedLm(dist ~ speed, data = cars)
```

Checking performance:
```{r 'Checking performance'}
cachedLm <- makeCached(lm)
system.time(cachedLm(dist ~ speed, data = cars))
system.time(cachedLm(dist ~ speed, data = cars))
```



